\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}

### 2022-04-05 lin-vs-cen.tex
Finally a full application of the LLL

The \defin{$r$-box} centred at a vertex $v:=(i,j)$ of $G_{k}$ is defined as
\[
  B_r(v) := \{i-r,\ldots,i+r\}\times\{j-r,\ldots,j+r\} \cap V(G_{k}) \enspace .
\]
The \defin{$r$-box} centred at an edge $vw$ of $G_{k}$ is $B_r(vw):=B_r(v)\cup B_r(w)$.
% Observe that, for any $\mu\in V(G_{k})\cup E(G_{a\times b})$,  $v\in B_r(u)$ if and only if $u\in B_r(v)$.
Any $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{k}[B_r(\mu)]$.  We extend these definitions to vertices of an $k\times k$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ as follows. For any $\mu\in V(G_k)$, define
\[
   \tilde{B}_r(\mu) := \bigcup_{\nu\in \VE(G_r(\mu))} V(P_\nu)
\]
and, for any $v\in V(G)$, let $\tilde{B}_r(v):=\tilde{B}_r(\mu_v)$
where $\mu_v$ denotes the unique element in $\VE(G_k)$ such that $v\in P_{\mu_v}$.\footnote{Technically, the notation for $\tilde{B}_r(v)$ should include the partition $\mathcal{P}$, but we omit this since there will never any ambiguity as to which partition is being used.}
% For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.

@@

  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\in \Omega(d/r^2)$ for each $\alpha\in X$.
  For each $\alpha\in X$, let $\Phi_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\Phi_\alpha| \ge d$.  Furthermore all but at most one element of $\Phi_\alpha$ is contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since, otherwise the first round would have succeeded for colour $\alpha$.
  % For each $\mu\in Y$, $|\VE(G_{r+1}(\mu))|\in O(r^2)$. Therefore, there are $\Omega(|\Phi_\alpha|/r^2)$ objects $\mu\in Y$ such that $\VE(G_{r+1}(\mu))\cap \Phi_\alpha\neq\emptyset$.

  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\Omega(|\Phi_\alpha|/r^2),1/25)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 100r^2)$.  (We make no attempt to optimize the constants here.)  Suppose that, for some $\mu\in Y$, $B_r(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$. Then, by \cref{packing_lemma}, the set
  \[
     Y_\nu := \{\tau\in Y:\nu\in\VE(G_{2r}(\tau))\}
  \]
  has size at most $25$. Observe that the edge $\alpha\mu$ will appear in $H$ if, in our random permutation $\pi$ of $Y$, $\nu$ appears before any other element of $Y_\nu$.  This happens with probability exactly $1/|Y_\nu|\ge 1/25$.
   % This already establishes that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$.

  In order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.  To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $8r+O(1)$ would be sufficient.}  This process continues until $B_r(\mu)\cap J=\emptyset$ for each $\mu\in Y$. Since $|\VE(G_{10r}(\mu))|\in O(r^2)$, each step in this process eliminates $O(r^2)$ objects from $J$ so the number $t$ of iterations is in $\Omega(|\Phi_\alpha|/r^2)\subseteq \Omega(d/r^2)$.

  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and let $\{\nu_1,\ldots,\nu_r\}\subseteq\Phi_\alpha$ be the corresponding elements of $\Phi_\alpha$.  The important observation now is that the sets $Y_{\nu_1},\ldots,Y_{\nu_t}$ are disjoint.  Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears in $\pi$ before any other element of $Y_{\nu_i}$'' then the events $U_1,\ldots,U_k$ are mutually independent.  Indeed, each $U_i$ depends only on the relative order of $Y_{\nu_i}$ in the random permutation of $Y$.  For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/25$.  Therefore, $\sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/25)$ random variable.  Therefore,
  \begin{align*}
    \Pr\left(\deg_H(\alpha)< 100r^2\right)
    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 100r^2\right) \\
    & \le \Pr(\binomial(t,1/25)\le 100r^2) \\
    & \le (2t)^{100r^2}(1/25)^{r^2}(1-1/25)^{t-100r^2} \\
    & \le \exp(100r^2\log(2t)+(t-100r^2)\log(1-1/?)) \\
    & \le \exp(-100r^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\
    & = \exp(ar^2 - bt)
  \end{align*}
  for fixed positive $a,b>0$.  Recall that we are working in the regime in  $t\in\Omega(|\Phi_\alpha|/r^2)$ and $|\Phi_\alpha|\in\Omega(d/r^2)$ so, for any fixed $C>0$, there exists $d\in\Omega(r^4)$ such that
  $\Pr(\deg_H(\alpha)< 100r^2)\le \exp(-C|\Phi_\alpha|/r^2)$.\todo{This could be done better.} Let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  By the preceding discussion, $\Pr(E_\alpha)\le \exp(-C|\Phi_\alpha|/r^2)$.

  We are now ready for an application of \cref{weighted_lovasz}.  Let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$ and, for each $\alpha\in X$, define
  \[
     \Gamma_\alpha : \left\{E_\beta \in \mathcal{E}: \Phi_\beta\cap \bigcup_{\nu\in \Phi_\alpha} \VE(G_{10r}(\nu))\neq\emptyset\right\} \enspace .
  \]
  The event $E_\alpha$ is mutually independent of $\mathcal{E}\setminus\Gamma_\alpha$ since, for any $E_\beta\in\mathcal{E}\setminus\Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\beta\neq\emptyset\}$ are disjoint and $\deg_H(\alpha)$ and $\deg_H(\beta)$ are each determined by the relative orders of $Y_\alpha$ and $Y_\beta$ in $\pi$, respectively.

  For each $\alpha\in X$, let $w(E_\alpha):=1-e^{-C/50r^2}$.  Then
  \begin{align}
    w(E_\alpha)\prod_{E_\beta\in\Gamma_\alpha} (1-w(E_\beta))
    & = (1-e^{-C/50r^2})e^{-C|\Gamma_\alpha|/50r^2} \\
    & = e^{-C|\Gamma_\alpha|/50r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2} \\
    & \ge e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2} \enspace . \label{ratioo}
    % & \ge (1-e^{-C/50r^2})e^{-C|\Phi_\alpha|/2r^2} \\
    % & \ge (1-e^{-C/50r^2})e^{-C|\Phi_\alpha|/2r^2} \\
  \end{align}
  Taking the ratio of \cref{ratioo} and $\Pr(E_\alpha)$ we obtain
  \begin{align*}
    \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{\Pr(E_\alpha)}
    & \ge \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
    & \ge e^{C|\Phi_\alpha|/2r^2} - \frac{e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
  \end{align*}

  \todo[inline]{Hmmmmm. Not working.  Try something like $w(E_\alpha)=1-e^{-1/50}$}

  Finally, by \cref{packing_lemma} $|\Gamma_\alpha|\le 25|\Phi_\alpha|$.

  \begin{align*}
      \Pr(E_\alpha) \le \exp(-|\Phi_\alpha|\cdot C/r^2) \enspace .
  \end{align*}

  For each $\alpha\in X$, define $w(E_\alpha)=C/2r^2$ and observe that
  \begin{align*}
    p(1-p)^{|\Gamma_\alpha|}
    & \ge  p(1-p)^{25|\Phi_\alpha|} = p\exp(|\Phi_\alpha|\cdot 25\log(1-c/2r^2))
  \end{align*}
  \todo[inline]{Sort out these calculations}

  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\alpha_1,\ldots,\alpha_{2|K|}$ such that $\deg_H(\alpha)\ge 100r^2$ for each $\alpha\in X$.  On the other hand, $\deg_{H}(y)\le 8r^2$ for each $y\in Y$.  Therefore, for any $A\subseteq X$, $|N_H(A)|\ge 100|A|/8 \ge 2|A|$.  Therefore, by \cref{d_hall}, there is a subgraph $M$ of $H$ in which $\deg_M(\alpha)=2$ for each $\alpha\in X$ and $\deg_M(\mu)\le 1$ for each $\mu\in Y$. Each edge $\alpha\mu$ of $M$ corresponds to some $\nu\in X_\alpha\cap\VE(G_r(\mu))$, which we place in $Q_2$.

@@

\begin{proof}[Proof of \cref{pseudogrid_lower_bound}]
  Suppose for the sake of contradiction that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours for arbitrarily small $\epsilon >0$.  Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)|\ge d|A|$, for any $d\le 1/2\epsilon$.  In particular, some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $B_r(\mu)\cap S = \{\mu\}$ for each $\mu\in S$.  By By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.  Since $P$ is contained in $G'$ $\varphi(V(P))=\varphi(V(G'))$, so $P$ is has no center under $\varphi$, contradicting the assumption that $\varphi$ is a linear colouring of $G$.
\end{proof}

>>>>

The \defin{$r$-box} centred at a vertex $v:=(i,j)$ of $G_{k}$ is defined as
\[
  B_r(v) := \{i-r,\ldots,i+r\}\times\{j-r,\ldots,j+r\} \cap V(G_{k}) \enspace .
\]
The \defin{$r$-box} centred at an edge $vw$ of $G_{k}$ is $B_r(vw):=B_r(v)\cup B_r(w)$.  
% Observe that, for any $\mu\in V(G_{k})\cup E(G_{a\times b})$,  $v\in B_r(u)$ if and only if $u\in B_r(v)$.
Any $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{k}[B_r(\mu)]$.  Straightforward counting shows that, for any $\mu\in\VE(G_k)$,
\begin{equation}
   |\VE(G_r(\mu))| \le 3(2r+1)^2 \enspace . \label{rbox_size} % rbox size
\end{equation}
We extend these definitions to vertices of a $k\times k$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ as follows. For any $\mu\in V(G_k)$, define
\[
   \tilde{B}_r(\mu) := \bigcup_{\nu\in \VE(G_r(\mu))} V(P_\nu)
\]
and, for any $v\in V(G)$, let $\tilde{B}_r(v):=\tilde{B}_r(\mu_v)$
where $\mu_v$ denotes the unique element in $\VE(G_k)$ such that $v\in P_{\mu_v}$.\footnote{Technically, the notation for $\tilde{B}_r(v)$ should include the partition $\mathcal{P}$, but we omit this since there will never any ambiguity as to which partition is being used.}
% For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.

@@

  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 6(2r+1)^2$ for each $\alpha\in X$.
  For each $\alpha\in X$, let $\Phi_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\Phi_\alpha| \ge d$.  Furthermore all but at most one element of $\Phi_\alpha$ is contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since, otherwise the first round would have succeeded for colour $\alpha$.
  % For each $\mu\in Y$, $|\VE(G_{r+1}(\mu))|\in O(r^2)$. Therefore, there are $\Omega(|\Phi_\alpha|/r^2)$ objects $\mu\in Y$ such that $\VE(G_{r+1}(\mu))\cap \Phi_\alpha\neq\emptyset$.

  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\ceil{|\Phi_\alpha|/3(20r+1)^2},1/25)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 6(2r+1)^2)$.  (We make no attempt to optimize the constants here.)  Suppose that, for some $\mu\in Y$, $B_r(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$. Then, by \cref{packing_lemma}, the set
  \[
     Y_\nu := \{\tau\in Y:\nu\in\VE(G_{2r}(\tau))\}
  \]
  has size at most $25$. Observe that the edge $\alpha\mu$ will appear in $H$ if, in our random permutation $\pi$ of $Y$, $\nu$ appears before any other element of $Y_\nu$.  This happens with probability exactly $1/|Y_\nu|\ge 1/25$.
   % This already establishes that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$.

  In order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.  To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $8r+O(1)$ would be sufficient.}  This process continues until $B_r(\mu)\cap J=\emptyset$ for each $\mu\in Y$. Since $|\VE(G_{10r}(\mu))|\le 3(20r+1)^2$, each step in this process eliminates at most $3(20r+1)^2$ objects from $J$ so the number $t$ of iterations is at least $\ceil{|\Phi_\alpha|/3(20r+1)^2}$

  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and let $\{\nu_1,\ldots,\nu_r\}\subseteq\Phi_\alpha$ be the corresponding elements of $\Phi_\alpha$.  The important observation now is that the sets $Y_{\nu_1},\ldots,Y_{\nu_t}$ are disjoint.  Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears in $\pi$ before any other element of $Y_{\nu_i}$'' then the events $U_1,\ldots,U_k$ are mutually independent.  Indeed, each $U_i$ depends only on the relative order of $Y_{\nu_i}$ in the  permutation $\pi$.  For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/25$.  Therefore, $\sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/25)$ random variable.\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.}  Therefore,
  \begin{align}
    \Pr\left(\deg_H(\alpha)< 6(2r+1)^2\right) 
    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 6(2r+1)^2\right) \notag \\
    & \le \Pr(\binomial(t,1/25)\le 6(2r+1)^2) \notag \\
    & \le (2t)^{6(2r+1)^2}(1/25)^{r^2}(1-1/25)^{t-6(2r+1)^2} \notag \\
    % & \le \exp(6(2r+1)^2\log(2t)+(t-6(2r+1)^2)\log(1-1/25)) \\
    & \le \exp(6(2r+1)^2(1+\log(2t))-t/25) \notag \\
    & \le \exp(6(2r+1)^2(1+\log(2|\Phi_\alpha|))-|\Phi_\alpha|/75(20r+1)^2) \label{probability}
    % & \le \exp(-6(2r+1)^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\
    % & = \exp(ar^2 - bt)
  \end{align}
  % for fixed positive $a,b>0$.  Recall that we are working in the regime in  $t\in\Omega(|\Phi_\alpha|/r^2)$ and $|\Phi_\alpha|\in\Omega(d/r^2)$ so, for any fixed $C>0$, there exists $d\in\Omega(r^4)$ such that
  % $\Pr(\deg_H(\alpha)< 6(2r+1)^2)\le \exp(-C|\Phi_\alpha|/r^2)$.\todo{This could be done better.}   By the preceding discussion, $\Pr(E_\alpha)\le \exp(-C|\Phi_\alpha|/r^2)$.

  We are now ready for an application of \cref{weighted_lovasz}.  For each $\alpha\in X$, let Let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 6(2r+1)^2$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$. For each $\alpha\in X$, define
  \[
     \Gamma_\alpha := \left\{E_\beta \in \mathcal{E}: \Phi_\beta\cap \bigcup_{\nu\in \Phi_\alpha} \VE(G_{10r}(\nu))\neq\emptyset\right\} 
  \]
  and observe that, by \ref{rbox_size}, $|\Gamma_\alpha|\le 3(20r+1)^2|\Phi_\alpha$.
  The event $E_\alpha$ is mutually independent of $\mathcal{E}\setminus\Gamma_\alpha$ since, for any $E_\beta\in\mathcal{E}\setminus\Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\beta\neq\emptyset\}$ are disjoint and $\deg_H(\alpha)$ and $\deg_H(\beta)$ are each determined by the relative orders of $Y_\alpha$ and $Y_\beta$ in $\pi$, respectively.

  For each $\alpha\in X$, let $w(E_\alpha):=1/\tau$, for some $\tau>0$ to be determined shortly.  Then
  \begin{align}
    w(E_\alpha)\prod_{E_\beta\in\Gamma_\alpha} (1-w(E_\beta))
    & = \tfrac{1}{\tau}\left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|-1} \notag \\
    & \ge \exp\left(-\log\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
    & \ge \exp\left(-\log\tau - 3(20r+1)^2|\Phi_\alpha|/(\tau-1)\right) \enspace . \label{weight_product}
  \end{align}

  Taking $\tau = 450(20r+1)^4+1$, and using \cref{weight_product,probability}, we find that
  \[
     \Pr\left(E_\alpha\right) \le w(E_\alpha)\prod_{\beta\in\Gamma_\alpha}(1-w(\Gamma_\alpha))
  \]
  provided that 
  \[   |\Phi_\alpha| \ge 150(20r+1)^2\left(\log\left(450(20r+1)^4+1\right) + 6(2r+1)^2(1+\log(2|\Phi_\alpha|)) \right) \enspace .
  \]
  The right hand side of this last equation is of the form $O(r^4\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d$, this can therefore be satisfied for some $d\in O(r^4)$.

  % Taking the ratio of \cref{ratioo} and $\Pr(E_\alpha)$ we obtain
  % \begin{align*}
  %   \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{\Pr(E_\alpha)}
  %   & \ge \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
  %   & \ge e^{C|\Phi_\alpha|/2r^2} - \frac{e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
  % \end{align*}
  % 
  % \todo[inline]{Hmmmmm. Not working.  Try something like $w(E_\alpha)=1-e^{-1/50}$}
  % 
  % Finally, by \cref{packing_lemma} $|\Gamma_\alpha|\le 25|\Phi_\alpha|$.
  % 
  % \begin{align*}
  %     \Pr(E_\alpha) \le \exp(-|\Phi_\alpha|\cdot C/r^2) \enspace .
  % \end{align*}
  % 
  % For each $\alpha\in X$, define $w(E_\alpha)=C/2r^2$ and observe that
  % \begin{align*}
  %   p(1-p)^{|\Gamma_\alpha|}
  %   & \ge  p(1-p)^{25|\Phi_\alpha|} = p\exp(|\Phi_\alpha|\cdot 25\log(1-c/2r^2))
  % \end{align*}
  % \todo[inline]{Sort out these calculations}

  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\alpha_1,\ldots,\alpha_{2|K|}$ such that $\deg_H(\alpha)\ge 6(2r+1)^2$ for each $\alpha\in X$.  On the other hand, $\deg_{H}(y)\le |\VE(G_r(y))|-1 \le 3(2r+1)^2$ for each $y\in Y$.  Therefore, for any $A\subseteq X$, $|N_H(A)|\ge 6(r+1)^2|A|/3(2r+1)^2 = 2|A|$.  Therefore, by \cref{d_hall}, there is a subgraph $M$ of $H$ in which $\deg_M(\alpha)=2$ for each $\alpha\in X$ and $\deg_M(\mu)\le 1$ for each $\mu\in Y$. Each edge $\alpha\mu$ of $M$ corresponds to some $\nu\in X_\alpha\cap\VE(G_r(\mu))$, which we place in $Q_2$.

@@

\begin{proof}[Proof of \cref{pseudogrid_lower_bound}]
  Suppose for the sake of contradiction that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours for arbitrarily small $\epsilon >0$.  Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)|\ge d|A|$, for any $d\le 1/2\epsilon$.  In particular, some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $B_r(\mu)\cap S = \{\mu\}$ for each $\mu\in S$.  By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.  Since $P$ is contained in $G'$ $\varphi(V(P))=\varphi(V(G'))$, so $P$ is has no center under $\varphi$, contradicting the assumption that $\varphi$ is a linear colouring of $G$.
\end{proof}
